Analyzing DECREASE trials to estimate evidence of data manipulation

### 2018-05-23 submission/review-reply.md

Thank you :)

> General Comments:

> Some circular reasoning in  step 2, In step 2 the authors provide clear evidence that the results of the DECREASE trials are quite unlikely, however, they employ some circular reasoning in their argumentation. Because we know the DECREASE values are extreme, when a p-value is calculated for the DECREASE effect is calculated based off of the non-DECREASE effect distribution it will of course be very small. The results of this approach are methodically sound but suffer from circular reasoning and do not seem to add much to the goal of determining whether data manipulation occurred. Although this step might be useful to help ease the reader into the following steps, the limited added value of these analyses should be recognized.

We recognize the circularity and adapt the text to indicate that this is a quantification of the suspicion that the data points are extreme. This quantification shows *how* extreme these data points are.

@@

> It would be helpful to have a forest plot of Table 1, to better visualize the results and compare DECREASE vs. non-DECREASE.

We did generate this forest plot at some point, and saw exactly the same one as in Bouri et al. (2014). We explicitly did not include a forest plot for that reason.

> In Table 2, the authors could report the relative risk (RR) instead of the log(RR) to improve the table’s readability and help the reader understand the magnitude of the different effects depicted. We would also be interested whether some more general information on this topic, especially regarding the non-DECREASE trials to put the DECREASE trials and their data into perspective. (e.g. how many non decrease trials were registered but not published; or whether there is any other evidence or suggestion for non-publication of relevant data that could alter some of the conclusions in this paper.)

We decided to stick to the log(RR) because otherwise readers would have to backtransform when trying to calculate the log(RR) for a specific set of conditions using the results. With respect to investigating how many trials had been registered, we agree it is an interesting research question on its own and consider a superficial investigation in the current paper to contextualize the findings not proportional to the complexity of the issue (e.g., a review of trial registration would be necessary given the first DECREASE trial took place in the late 90s). Hence, we do not take up this suggestion here.

@@

> We suggest rewording the second paragraph under the results section, and specifically address the sentence that begins “The relatively minor difference…”. This sentence first references “estimates”, but it is unclear what these estimates are referring to.

Adjusted for clarity to "The relatively minor difference between the point- and distribution estimates indicates "

> In Figure 2 please address why, in the bottom lower panel, the confidence interval for the point estimate is larger than the distribution estimate. Because the distribution estimate has more error in it, we were expecting that the confidence interval of the distribution estimate would be larger than that of the point estimate.

We also found this odd and checked the code multiple times to find any mistakes. We invite anyone to take a look and correct us if we're wrong, but we are unable to find the mistake if there is one. Notably, it is the same code that gives the results for DECREASE-I that do follow the notion that the point estimates are more precise. We added a sentence to make this explicit to the reader: "We note that, for reasons unknown to us, the confidence interval of the point estimates are wider than those of the distribution estimates, whereas similar code for DECREASE-I yield the reverse (the code is available for review at https://osf.io/4zs29/)."

### 2018-05-23 submission/manuscript.Rmd
Update pre-revisions

For DECREASE-IV ($N=`r bouri_data$control_n[11] + bouri_data$beta_n[11]`$), the 95\% confidence interval for the estimated number of manipulated data points is [`r which.min(res_iv_certain - .025) - 1` - `r which.max(res_iv_certain - .975) - 1`] or [`r which.min(res_iv_uncertain - .025) - 1` - `r which.max(res_iv_uncertain - .975) - 1`] when based on a point estimate or a more uncertain distribution estimate, respectively. The relatively minor difference between the point- and distribution estimates indicates that there is a high degree of confidence that data manipulation did occur based on the difference of the trial results alone. Nonetheless, the range of potentially manipulated data points is still estimated at approximately 1000; this indicates that the summary results are insufficient to provide more than an estimated lower bound. This indicates that it is possible not all data were manipulated (i.e., $N=`r length(res_iv_certain) - 1`$) but at least some were, increasing the importance of well-documented data provenance to discern between genuine and falsified data. We note that, for reasons unknown to us, the confidence interval of the point estimates are more uncertain than those of the distribution estimates, whereas similar code for DECREASE-I provides the reverse (the code is available for review at https://osf.io/4zs29/).

@@

The results of our analyses also highlight that, despite the lack of availability of the raw data, summary results from larger samples allow for marginally better estimates of the number of manipulated data points when similar trials are available. Larger trials result in relatively more certainty (e.g., DECREASE-IV) about the estimated number of manipulated data points, when using the inversion method, compared to smaller trials (e.g., DECREASE-I). This increased certainty is due to decreased standard errors of the estimated effects, resulting in higher sensitivity to data anomalies. Nonetheless, much residual uncertainty remains and simply less information is available in summary results when compared to raw data, as indicated by the large range of the confidence intervals in Step 3. As such, raw data availability would improve the options open to detect potential anomalies (note: raw data are available for DECREASE VI, but upon a freedom of information request by the first author, Erasmus MC refused to share these data; see [osf.io/fswjt/](https://osf.io/fswjt/)). The results also highlight that in order to prevent detection, it would be in the manipulator's interest to fabricate small and imprecise studies (assuming the manipulator wants to remain undetected), which ultimately detracts from the scientific value of such a study and hence the individual reward for manipulation through reduced impact (hopefully).

>>>>

For DECREASE-IV ($N=`r bouri_data$control_n[11] + bouri_data$beta_n[11]`$), the 95\% confidence interval for the estimated number of manipulated data points is [`r which.min(res_iv_certain - .025) - 1` - `r which.max(res_iv_certain - .975) - 1`] or [`r which.min(res_iv_uncertain - .025) - 1` - `r which.max(res_iv_uncertain - .975) - 1`] when based on a point estimate or a more uncertain distribution estimate, respectively. The relatively minor difference between the point- and distribution estimates indicates that there is a high degree of confidence that data manipulation did occur based on the difference of the trial results alone. Nonetheless, the range of potentially manipulated data points is still estimated at approximately 1000; this indicates that the summary results are insufficient to provide more than an estimated lower bound. This indicates that it is possible not all data were manipulated (i.e., $N=`r length(res_iv_certain) - 1`$) but at least some were, increasing the importance of well-documented data provenance to discern between genuine and falsified data. We note that, for reasons unknown to us, the confidence interval of the point estimates are wider than those of the distribution estimates, whereas similar code for DECREASE-I yield the reverse (the code is available for review at https://osf.io/4zs29/).

@@

The results of our analyses also highlight that, despite the lack of availability of the raw data, summary results from larger samples allow for marginally better estimates of the number of manipulated data points when similar trials are available. Larger trials result in relatively more certainty (e.g., DECREASE-IV) about the estimated number of manipulated data points, when using the inversion method, compared to smaller trials (e.g., DECREASE-I). This increased certainty is due to decreased standard errors of the estimated effects, resulting in higher sensitivity to data anomalies. Nonetheless, much residual uncertainty remains and simply less information is available in summary results than in raw data, as indicated by the width of the confidence intervals in Step 3. As such, raw data availability would improve the options open to detect potential anomalies (note: raw data are available for DECREASE VI, but upon a freedom of information request by the first author, Erasmus MC refused to share these data; see [osf.io/fswjt/](https://osf.io/fswjt/)). The results also highlight that in order to prevent detection, it would be in the manipulator's interest to fabricate small and imprecise studies (assuming the manipulator wants to remain undetected), which ultimately detracts from the scientific value of such a study and hence the individual reward for manipulation through reduced impact (hopefully).
